---
title: "Yearning for Stationarity"
author: "Philip Anderson; panders2@tamu.edu"
date: "6/10/2018"
output: pdf_document
---

```{r setup, message=FALSE}
library("astsa")
library("DescTools")
library("GGally")
library("tseries")
```

Objective: I'm going to mess around with some of the newer stuff we have covered.  Also going to build on some of the regression work that Colby shared and see if I can add anything there.

```{r}
# read in the data
mri <- read.csv("/Users/panders2/Documents/schools/tamu/stat_626/project/stat_626_proj/mri_dat_one.csv")
# update field names
names(mri) <- c("hour", "minute", "freq", "int_pressure", "atm_pressure", "tot_pressure", "tesla")
```

With an initial look at the data, I think we need to smooth it to eliminate the abberration at time ~950.  Let's take a look.

```{r}
astsa::tsplot(mri$freq)
```

Zoom in:

```{r}
astsa::tsplot(mri$freq[900:1000])
```

I don't know how much of an impact this outlier will have on any sort of models we build, but it probably isn't good.  Regardless, taking care of it will be a good exercise in smoothing, so let's do it anyway.  The _ksmooth_ function used by the book does not appear to have automatic bandwidth selection, so let's try several and see what happens.

```{r}

astsa::tsplot(mri$freq)
lines(ksmooth(time(mri$freq), mri$freq, kernel="normal", bandwidth=0.5), lwd=1.5, col="blue")
lines(ksmooth(time(mri$freq), mri$freq, kernel="normal", bandwidth=5), lwd=1.5, col="red")
lines(ksmooth(time(mri$freq), mri$freq, kernel="normal", bandwidth=15), lwd=1.5, col="green")

legend("bottomleft", c("bw=0.5", "bw=5", "bw=15"), lwd=rep(1.5, 3) , col=c("blue", "red", "green"))

```

Focusing on just the problem area, it seems that a higher bandwidth does the best job at smoothing this area out.  There may be some concern about over-smoothing, but we can proceed as is.  Once again, let's zoom in.

```{r}
# (I believe) kernel smoothing utilizes the entire array for producing estimates, so we will 
# need to make new variables on the entire set before zooming

mri$freq_point5 <- ksmooth(time(mri$freq), mri$freq, kernel="normal", bandwidth=0.5)$y
mri$freq_5 <- ksmooth(time(mri$freq), mri$freq, kernel="normal", bandwidth=5)$y
mri$freq_15 <- ksmooth(time(mri$freq), mri$freq, kernel="normal", bandwidth=15)$y

astsa::tsplot(mri$freq[900:1000], lwd=2)
lines(mri$freq_point5[900:1000], lwd=1.25, col="blue")
lines(mri$freq_5[900:1000], lwd=1.25, col="red")
lines(mri$freq_15[900:1000], lwd=1.25, col="green")

legend("bottomleft", c("bw=0.5", "bw=5", "bw=15"), lwd=rep(1.5, 3) , col=c("blue", "red", "green"))
```

Kernel smoothing is one viable option - localized regression is another.  Let's do that and compare.  Like kernel bandwidth parameter, the lowess regressor incorporates a parameter that influences smoothness, (_f_).  It simply measures what proportion of the array points we should use to influence the smooth at each estimation point.

```{r}
astsa::tsplot(mri$freq, lwd=2)
lines(lowess(mri$freq, f=0.1), lwd=1.25, col="blue")
lines(lowess(mri$freq, f=0.2), lwd=1.25, col="purple")
lines(lowess(mri$freq, f=0.5), lwd=1.25, col="red")
lines(lowess(mri$freq, f=1), lwd=1.25, col="green")

legend("bottomleft", c("f=0.1", "f=0.2", "f=0.5", "f=1"), lwd=rep(1.5, 5) , col=c("blue", "purple",  "red", "green"))
```

The kernel smoother hugs the original trend line much more closely than the lowess regression.  I am going to add the f=0.1 and f=0.2 estimations to the data, because they seem more fair than 0.5 or 1.


```{r}
mri$freq_low_1 <- lowess(mri$freq, f=0.1)$y
mri$freq_low_2 <- lowess(mri$freq, f=0.2)$y

```

We now have a handful of smoothed data points.  Let's see if we can get to stationarity using those as our base.  Taking differences seemed promising in the past so let's do that again.

```{r}
astsa::tsplot(diff(mri$freq))
```

This looks stationary, but it may just be an axis thing.  Let's adjust it and revisit. 

```{r}
astsa::tsplot(diff(mri$freq), ylim=c(-1.5,1.5))
```

This is not looking quite as stationary as before, which we have already seen the empirical ACF will corroborate.

```{r}
acf(diff(mri$freq))
```

It's tough to see in the differenced time series plot, but we do have a strong and significant negative lag at lag 1.

Let's repeat for some of our smoothed series and see.

```{r}
# kernel smoother with bandwidth of 15
astsa::tsplot(diff(mri$freq_15))
```

Not good.

```{r}
acf(diff(mri$freq_15))
```

Let's just try the simpler moving average instead of fancy things.

```{r}
# 6 point MA
mri$freq_ma <- stats::filter(mri$freq, sides=2, filter=rep(1/12, 12))
astsa::tsplot(diff(mri$freq_ma))
```

Bleh.

Ironic, but let's add noise to the smoothed series and see if that helps.

```{r}

test <- mri$freq_low_1 + rnorm(length(mri$freq_point5))
astsa::tsplot(test)
astsa::tsplot(diff(test))
acf(diff(test))
```

That was promising for a brief second.

```{r}
lin_mod <- lm(mri$freq ~ 0 + mri$atm_pressure + mri$int_pressure)
astsa::tsplot(diff(lin_mod$residuals))
acf(diff(lin_mod$residuals))
```

```{r}
lin_mod2 <- lm(mri$freq_point5 ~ time(mri$freq_point5))
astsa::tsplot(diff(lin_mod2$residuals))
acf(diff(lin_mod2$residuals, differences=2))

```


```{r}
adf.test(diff(mri$freq))
kpss.test(diff(mri$freq))
```


```{r}
acf(diff(mri$freq[1:800]))
acf(diff(mri$freq[801:1600]))

```


I have found structural modeling to be the most interesting thing covered in this course so far.  I have found that the _decompose_ function requires evidence of periodicity that we do not have.

```{r, error=F}
decompose(mri$freq_low_2)
```



```{r}

```


```{r}
astsa::tsplot(mri$freq)
astsa::tsplot(log(mri$freq))
astsa::tsplot(diff(log(mri$freq), lag=1))

```


```{r}
head(mri)
```

```{r}
GGally::ggpairs(mri[, c(3,4,5)])
```


```{r}
# regression of freq on time

lin_mod_1 <- lm(mri$freq ~ time(mri$freq))
summary(lin_mod_1)
plot(lin_mod_1)
```

```{r}

semi_mod <- mgcv::gam(mri$freq ~ s(time(mri$freq)))
plot(semi_mod)
summary(semi_mod)

```

```{r}
ts.plot(mri$freq[940:960])

```

Let's kernel-smooth that abberration out of there.

```{r}
astsa::tsplot(mri$freq)
lines(ksmooth(time(mri$freq), mri$freq, "normal", bandwidth=10), lwd=2, col=4)
#lines(ksmooth(time(soi), soi, "normal", bandwidth=1), lwd=2, col=4)
```

```{r}
freq_reduced <- mri$freq[900:1100]
ts.plot(freq_reduced)
lines(ksmooth(time(freq_reduced), freq_reduced, "normal", bandwidth=5), lwd=1.5, col="blue") 
```

```{r}
#mri$smooth_freq <- ksmooth(mri$freq, )

mri$smooth_freq <- ksmooth(time(mri$freq), mri$freq, kerne="normal", bandwidth=15)$y
```

```{r}
astsa::tsplot(mri$freq)
astsa::tsplot(mri$smooth_freq)
```

```{r}
astsa::tsplot(mri$smooth_freq)
astsa::tsplot(diff(mri$smooth_freq))

acf(diff(mri$smooth_freq))

```


```{r}

linear <- lm(mri$smooth_freq ~ mri$int_pressure + mri$atm_pressure)
summary(linear)
plot(linear)
```


```{r}
ccf.Freq.IntPrs = ccf(mri$freq, mri$int_pressure)
```

```{r}
#index = which(ccf.Freq.IntPrs$acf == max(ccf.Freq.IntPrs$acf))
which(ccf.Freq.IntPrs$acf == max(ccf.Freq.IntPrs$acf))
```

```{r}
#lagData = ts.intersect(cbind(Frequency,AtmPrss),lagIntPrss = lag(IntPrss,sel.lag))



```


```{r}
astsa::tsplot(mri$freq)
lines(lowess(mri$freq, f=0.2))

```


```{r}
mini <- data.frame(
    sf_diff=diff(mri$smooth_freq)
    , int_diff=diff(mri$int_pressure)
    , atm_diff=diff(mri$atm_pressure)
    )
GGally::ggpairs(mini)

```

```{r}
lag2.plot(mri$smooth_freq, mri$atm_pressure, 8)
```




```{r}
Frequency <- mri$tesla
AtmPrss <- mri$atm_pressure
IntPrss <- mri$int_pressure

n = min(which(IntPrss == min(IntPrss)))-1
Frequency = Frequency[1:n]
AtmPrss = AtmPrss[1:n]
IntPrss = IntPrss[1:n]

```



```{r}
(n <- min(which(IntPrss == min(IntPrss))) - 1)
Frequency = Frequency[1:n]
AtmPrss = AtmPrss[1:n]
IntPrss = IntPrss[1:n]



```


```{r}

ccf(Frequency, IntPrss)
ccf(Frequency, AtmPrss)
```




