---
title: "STAT626_MRI"
author: "Philip Anderson; panders2@tamu.edu"
date: "6/3/2018"
output: pdf_document
---

```{r setup, message=FALSE}
library("tidyverse")
library("astsa")
```

## Objective:
I am doing a freestyle EDA process to explore the data and get myself and everyone else more acquainted with it.  I'll focus on the techniques we have covered so far.

Import the data

```{r}
# read in the data
mri <- read.csv("/Users/panders2/Documents/schools/tamu/stat_626/project/stat_626_proj/mri_dat_one.csv")
# update fields
names(mri) <- c("hour", "minute", "freq", "int_pressure", "atm_pressure", "tot_pressure", "tesla")
# take a look
str(mri)
head(mri)
summary(mri)
```

## Remaking Eric's Plots
### Primary Plot First

```{r}
plot(mri$freq, col="firebrick4"
    , main="Magnetic Field - Frequency vs. Time"
    , xlab="Time"
    , ylab="Frequency"
     )
mtext("Primary Variable of Interest")
```

### Covariate Plots Second

```{r}
plot(mri$atm_pressure, col="darkgreen", type="b"
    , main="Atmospheric Pressure vs. Time"
    , xlab="Time"
    , ylab="Frequency"
     )
mtext("Secondary Variable of Interest")
```


```{r}
plot(mri$int_pressure[1:1600], col="darkblue", type="b"
    , main="Controlled Internal Pressure vs. Time"
    , xlab="Time"
    , ylab="Frequency"
     )
mtext("Secondary Variable of Interest")
```

### Empirical Autocorrelation Functions

```{r}
freq_acf <- acf(mri$freq, lag.max=50
                , main="ACF for Magnetic Frequency" 
                )
```

So our main variable has a LOT of serial correlation and is far from stationary.  This could be because of time iteration.  Consider aggregation to every five minutes as Eric mentioned.

ACF for Secondary Variables

```{r}
acf(mri$atm_pressure, lag.max=50
    , main="ACF for Atmospheric Pressure"
    )
```

Same story as above.

```{r}
acf(mri$int_pressure[1:1600], lag.max=100
   , main="ACF - Internal Pressure" 
  )
    
```

ACF for Internal Pressure variable is interesting - reflects the cyclic nature of series.

## Primary Series Manipulations

### Detrending + First Differences

```{r}
par(mfrow=c(3,1))
# plot 1
astsa::tsplot(mri$freq, main="Original Frequency Data")

# plot 2
detrend_fit <- lm(mri$freq ~ time(mri$freq))
detrend_series <- resid(detrend_fit)
astsa::tsplot(detrend_series, main="Detrended Series")

# plot 3
diff_series <- diff(mri$freq)
astsa::tsplot(diff_series, main="Differenced Series")
```

The strange point in the middle may be influencing the images.  It may be important, but lets take it out and try again.

```{r}
print(length(mri$freq))
# take away minimum value
mri_freq2 <- mri$freq[-(which.min(mri$freq))]
# make sure that worked
print(length(mri_freq2))
```

```{r}
par(mfrow=c(3,1))
# plot 1
astsa::tsplot(mri_freq2, main="Original Frequency Data")

# plot 2
detrend_fit <- lm(mri_freq2 ~ time(mri_freq2))
detrend_series <- resid(detrend_fit)
astsa::tsplot(detrend_series, main="Detrended Series")

# plot 3
diff_series <- diff(mri_freq2)
astsa::tsplot(diff_series, main="Differenced Series")
```

That's somewhat better.  Detrending did nothing but change the axis.  Differencing looks stationary?

```{r}
astsa::tsplot(diff_series, main="Differenced Series", col="firebrick4")
```



```{r}
acf(diff_series, main="ACF for Differenced Series")
```

Hard to see, but the ACF at a lot of these lag values is significant.  We are, however, much closer to stationarity than we were. 

## Formal Stationarity Check 

```{r}
Box.test(mri$freq, type="Ljung-Box")
Box.test(detrend_series, type="Ljung-Box")
Box.test(diff_series, type="Ljung-Box")
```

My understanding of this procedure is that we are looking for a high p-value.  We are consistently rejecting the null that the data are independently distributed (https://en.wikipedia.org/wiki/Ljung–Box_test).

```{r}
x <- rnorm(500, 0 ,1 )
Box.test(x, type="Ljung-Box")
```

## Cross-Correlation Functions

```{r}
ccf(mri$freq, mri$atm_pressure
    , main="CCF between Magnetic Frequency and Atmospheric Pressure"
      )
```

Yikes.

```{r}
ccf(mri$freq, mri$int_pressure, lag.max=50
    , main="CCF between Magnetic Frequency and Internal Pressure"
    )
```

Also yikes, but less so.

Let's swap out the main series for the differenced one.

```{r}
ccf(diff_series, mri$atm_pressure
    , main="CCF between Differenced Series and Atmospheric Pressure"
    )
```

```{r}
ccf(diff_series, mri$int_pressure
    , main="CCF between Differenced Series and Internal Pressure"
    )
```

This looks promising.

