---
title: "mri_dat_three_analysis"
author: "Philip Anderson; panders2@tamu.edu"
date: "6/17/2018"
output: pdf_document
---

# Objective
The purpose of this analysis is to try out some regressions of the data.

```{r setup, include=FALSE}
library("astsa")
library("GGally")
library("tidyverse")
library("mgcv")
```

```{r}
# read in the data
mri <- read.csv("/Users/panders2/Documents/schools/tamu/stat_626/project/stat_626_proj/mri_dat_one.csv")
# update field names
names(mri) <- c("hour", "minute", "freq", "int_pressure", "atm_pressure", "tot_pressure", "tesla")
```

First, replace the issue values with their imputations.

```{r}
tsplot(mri$freq[940:950])
```

The problems are the points at index 945.

```{r}
mri$freq[944] <- ((mri$freq[943] + mri$freq[945]) / 2)
```

```{r}
astsa::tsplot(mri$freq)
```

That's better.

```{r}
# parwise scatter plots of our data
GGally::ggpairs(mri%>%dplyr::select("freq", "int_pressure", "atm_pressure"))
```

These are not particularly helpful, probably because of the time series nature of the data.

Fit naive linear regressions

```{r}
basic <- lm(mri$freq ~ time(mri$freq))
plot(density(resid(basic)))
astsa::tsplot(resid(basic))
```

Errors are obviously serially correlated.  Now try this with the differenced series to see if it makes a difference.

```{r}

diff_mod <- lm(diff(mri$freq) ~ time(mri$freq[2:length(mri$freq)]))

plot(density(resid(diff_mod)))
astsa::tsplot(resid(diff_mod))
```

Looks better.

The ACF+PACF of the differenced series suggest an MA(1) model; let's try fitting this.

```{r}
serial_mod <- sarima(
                    xdata=diff(mri$freq)
                    , p=c(0)
                    , d=c(0)
                    , q=c(1)
                    , xreg=c(time(mri$freq[2:length(mri$freq)]))
                    )
```
Looks like there are some problems.  

Try semi-parametric regression (Generalized Additive Model) to see if we can get some explanation for the strange-looking scatterplot.  I'm not sure if there is a specialized version of these for time series, and we haven't covered them, but they might be somewhat helpful.

```{r}
semi_mod <- mgcv::gam(freq ~ s(int_pressure) + s(atm_pressure), data=mri)
plot(semi_mod, shade=T)
```


```{r}
semi_mod2 <- mgcv::gam(diff(freq) ~ s(int_pressure[2:length(int_pressure)]) + 
                         s(atm_pressure[2:length(atm_pressure)])
                       , data=mri)
plot(semi_mod, shade=T)
```

The graphics are showing the individual effects of each of the predictors on the outcome of _frequency_, controlling for other predictors.  The bottom of each graphic has a "rug" that will show you where the data points lie.  

In both cases, the effect of _internal_pressure_ is unclear - most of the points are centered around a few small values.  Atmospheric pressure seems to tell a more interesting story - as we increase the value of it, its effects on the predictor increase.




